---
title: "Statistical Modeling with Texas Housing Dataset"
output:
  pdf_document: default
  word_document: default
---

```{r}
# Set CRAN mirror
options(repos = c(CRAN = "https://cran.rstudio.com"))

# Install and load necessary packages
install.packages("ggplot2")
library(ggplot2)

# Load dataset
data("txhousing", package = "ggplot2")
```


Exploring the Dataset

```{r}
txhousing_class <- class(txhousing)
```

```{r}
colnames(txhousing)
```

```{r}
txhousing_dimensions <- dim(txhousing)

class_and_dimensions <- c(txhousing_class, txhousing_dimensions)

class_and_dimensions

```

```{r}
ls.str(txhousing)

```

```{r}
median(txhousing$sales, na.rm = TRUE)

```

```{r}
year_table <- table(txhousing$year)

year_table
```

```{r}
city_table <- table(txhousing$city)

dim(city_table)

```

Visualizing Variables of Interest

```{r}
boxplot(txhousing$sales, horizontal = TRUE, 
        main = "Sales of Texas Housing", 
        xlab = "Sales")

```

```{r}
boxplot(sales~year, 
        data = txhousing,
        horizontal = TRUE, 
        main = "Sales of Texas Housing over years", 
        xlab = "Sales",
        ylab = "Years",
        bty = "n",
        las = 1)

#lower sales in early 2000, sales between 0-2000 people was higher in 2008ish, fewer sales 2015.

```

```{r}

boxplot(median~year, 
        data = txhousing,
        horizontal = TRUE, 
        main = "Median Sales Price of Texas Housing Over Years", 
        xlab = "Median sales price",
        ylab = "Years",
        bty = "n",
        las = 1,
        col = "lightgreen")


boxplot(sales~year, 
        data = txhousing,
        horizontal = TRUE, 
        main = "Sales of Texas Housing Over Years", 
        xlab = "Sales",
        ylab = "Years",
        bty = "n",
        las = 1)


# Potentially, given the 2008 financial crisis, as housing prices maintained and even increased 
# around 2008, it is possible that the crisis has affected peoples ability to buy houses
# or at the least, the sales of houses in Texas.
```

```{r}
scatter.smooth(x = txhousing$median, 
               y = txhousing$sales, 
               main = "Sales Comparied to Median Sale Price", 
               xlab = "Sales",
               ylab = "Median sales price",
               cex = 0.75)

#Sales (independent = x) is the consequence of median sale price (dependent = y)


#The form is linear with clusters above correlated line
#The strength of sales or correlation seems tied to lower prices.
#The direction of the association has an upward trend but disperses after the 220000 mark.
```

Cleaning—Identifying and Removing NAs

```{r}
# Number of missing rows.

sum(is.na(txhousing$median))
```

```{r}
# Proportion of missing rows.

mean(is.na(txhousing$median))
```

```{r}
# Indices of missing rows.

NA_median_index <- which(is.na(txhousing$median))

```

```{r}
# Dataframe of rows with missing values (selected columns).

txhousing[NA_median_index, c("city", "year", "sales", "median")]
```

```{r}
# Cleaned dataframe (txhousing_clean).

txhousing_clean <- txhousing[-NA_median_index, c("city", "year", "sales", "median", "month")]

```

Modeling

```{r}
# Linear model mod1.
# Linear model: median sales price vs. sales.
mod1 <- lm(median ~ sales, data = txhousing_clean)

# Summary of the Linear Model.
summary_mod1 <- summary(mod1)
summary_mod1

# Check if sales is statistically significant (p-value < 0.05)
is_significant <- summary_mod1$coefficients["sales", "Pr(>|t|)"] < 0.05
is_significant

# Correlation between Sales and Median.
correlation <- round(cor(txhousing_clean$sales, txhousing_clean$median, use = "complete.obs"), 2)
correlation


# This model adds on to mod1 by adding an additional predictor, month, to see if
# both sales and month together predict median.
# Multiple regression Model: median sales price vs. sales and month
mod2 <- lm(median ~ sales + month, data = txhousing_clean)

# Summary of the multiple regression model
summary_mod2 <- summary(mod2)
summary_mod2

# Checking if sales and month are statistically significant
sales_significant <- summary_mod2$coefficients["sales", "Pr(>|t|)"] < 0.05
month_significant <- summary_mod2$coefficients["month", "Pr(>|t|)"] < 0.05

# R² value for mod2 and its interpretation
r_squared_mod2 <- summary(mod2)$r.squared
r_squared_mod2
# Interpretation of above:
# R² indicates the proportion of variance in 'median' explained by 'sales' and 'month'.
